---
title: "Introduction to fb4package"
author: "Hans Ttito"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to fb4package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 7,
  fig.height = 5
)
library(fb4package)
```

## Overview

`fb4package` is a programmatic R implementation of Fish Bioenergetics 4.0
(Deslauriers et al. 2017). It partitions consumed energy into metabolism,
wastes, and growth, providing mechanistic estimates of fish consumption and
growth under varying environmental conditions.

The main workflow has three steps:

1. **Build** a `Bioenergetic` object with species parameters, temperature,
   diet, and initial conditions.
2. **Run** `run_fb4()` to simulate or fit a bioenergetics model.
3. **Analyse / visualise** the results.

---

## 1. Species parameters

The package ships with a built-in database covering 105 models (73 species).

```{r species-db}
data(fish4_parameters)
cat("Species in database:", length(fish4_parameters), "\n")
head(names(fish4_parameters), 8)
```

Retrieve Chinook salmon (*Oncorhynchus tshawytscha*) and inspect one life
stage:

```{r chinook-params}
chinook_db  <- fish4_parameters[["Oncorhynchus tshawytscha"]]
stage       <- names(chinook_db$life_stages)[1]
cat("Life stage used:", stage, "\n")

sp_params   <- chinook_db$life_stages[[stage]]
sp_info     <- chinook_db$species_info

# Key consumption parameters
cat("CEQ:", sp_params$consumption$CEQ,
    "  CA:", sp_params$consumption$CA,
    "  CB:", sp_params$consumption$CB, "\n")
```

---

## 2. Environmental and diet data

Temperature and diet are supplied as data frames with a `Day` column.

```{r env-diet}
days <- 1:100

# Seasonal temperature (°C)
temp_data <- data.frame(
  Day         = days,
  Temperature = round(10 + 5 * sin(2 * pi * (days - 60) / 365), 2)
)

# Two-prey diet (constant proportions, J/g energy densities)
diet_data <- list(
  proportions = data.frame(Day = days, Alewife = 0.7, Shrimp = 0.3),
  prey_names  = c("Alewife", "Shrimp"),
  energies    = data.frame(Day = days, Alewife = 4900, Shrimp = 3200)
)

head(temp_data, 3)
```

---

## 3. Building a `Bioenergetic` object

```{r bio-object}
bio <- Bioenergetic(
  species_params      = sp_params,
  species_info        = sp_info,
  environmental_data  = list(temperature = temp_data),
  diet_data           = diet_data,
  simulation_settings = list(initial_weight = 1800, duration = 100)
)

# Required when PREDEDEQ = 2 (energy density as function of weight)
bio$species_params$predator$ED_ini <- 4500
bio$species_params$predator$ED_end <- 5500

print(bio)
```

---

## 4. Running simulations

### 4a. Direct simulation — fixed feeding level

Run the model with a constant proportion of maximum consumption (*p* = 0.8):

```{r direct}
res_direct <- run_fb4(bio, strategy = "direct", p_value = 0.8)
print(res_direct)
```

### 4b. Binary search — fit to a target final weight

Find the *p* that produces a target weight of 3 000 g after 100 days:

```{r binary-search}
res_bs <- run_fb4(bio,
                  strategy  = "binary_search",
                  fit_to    = "Weight",
                  fit_value = 3000)

cat("Estimated p-value :", round(res_bs$summary$p_value, 4), "\n")
cat("Final weight      :", round(res_bs$summary$final_weight, 1), "g\n")
cat("Total consumption :", round(res_bs$summary$total_consumption_g, 1), "g\n")
```

### 4c. MLE — uncertainty on *p* from observed weights

When multiple fish weights are available, `strategy = "mle"` returns a
point estimate of *p* together with its standard error and 95% confidence
interval:

```{r mle}
set.seed(42)
obs_weights <- rnorm(30, mean = 3000, sd = 80)   # 30 observed final weights

res_mle <- run_fb4(bio,
                   strategy         = "mle",
                   fit_to           = "Weight",
                   observed_weights = obs_weights,
                   upper            = 1.5)

cat("p-value (MLE)  :", round(res_mle$summary$p_value, 4), "\n")
cat("SE             :", round(res_mle$summary$p_se,    4), "\n")
cat("Converged      :", res_mle$summary$converged, "\n")
```

### 4d. Bootstrap — non-parametric uncertainty

```{r bootstrap}
set.seed(42)
res_boot <- run_fb4(bio,
                    strategy         = "bootstrap",
                    fit_to           = "Weight",
                    observed_weights = obs_weights,
                    n_bootstrap      = 200,
                    upper            = 1.5,
                    parallel         = FALSE)

cat("p mean (bootstrap):", round(res_boot$summary$p_mean, 4), "\n")
cat("p SD              :", round(res_boot$summary$p_sd,   4), "\n")
```

---

## 5. Comparing strategies

A quick look at how each method estimates *p* for the same data:

```{r compare, echo=FALSE}
methods <- c("binary_search", "mle", "bootstrap")
p_vals  <- c(
  res_bs$summary$p_value,
  res_mle$summary$p_value,
  res_boot$summary$p_mean
)
cat(sprintf("%-15s  p = %.4f\n", methods, p_vals))
```

---

## 6. Analysing results

```{r analysis}
growth   <- analyze_growth_patterns(res_bs)
feeding  <- analyze_feeding_performance(res_bs)
energy   <- analyze_energy_budget(res_bs)

cat(sprintf("Final weight              : %.1f g\n",
            growth$final_weight$estimate))
cat(sprintf("Specific growth rate      : %.4f g/g/day\n",
            growth$specific_growth_rate$estimate))
cat(sprintf("Total consumption         : %.1f g\n",
            feeding$total_consumption$estimate))
cat(sprintf("Gross conversion efficiency: %.3f\n",
            feeding$gross_conversion_efficiency$estimate))
```

---

## 7. Visualising results

```{r plot-growth, fig.cap="Daily growth trajectory from binary search fit."}
plot(res_bs, type = "growth")
```

```{r plot-energy, fig.cap="Daily energy budget partitioning."}
plot(res_bs, type = "energy")
```

```{r plot-dashboard, fig.cap="Full dashboard (growth, consumption, temperature, energy)."}
plot(res_bs, type = "dashboard")
```

---

## 8. Temperature × p sensitivity

Explore how final weight varies across temperatures and feeding levels:

```{r sensitivity, fig.cap="Temperature and p-value sensitivity grid."}
plot(bio,
     type         = "sensitivity",
     temperatures = seq(6, 18, by = 3),
     p_values     = seq(0.4, 1.0, by = 0.1))
```

---

## Energy balance equation

The model follows:

$$G = C - (R + F + U + \text{SDA})$$

where **G** = growth, **C** = consumption, **R** = respiration,
**F** = egestion, **U** = excretion, and **SDA** = specific dynamic action.
Consumption is:

$$C = C_{\max}(W,T) \cdot p$$

with $C_{\max} = C_A \cdot W^{C_B}$ scaled by a temperature function
$f(T)$ specific to each species equation (CEQ 1–4).

---

## References

Deslauriers D, Chipps SR, Breck JE, Rice JA, Madenjian CP (2017). Fish
Bioenergetics 4.0: An R-Based Modeling Application. *Fisheries* 42(11):586–596.
<https://doi.org/10.1080/03632415.2017.1377558>

Hewett SW, Johnson BL (1992). Fish Bioenergetics Model 2. University of
Wisconsin Sea Grant Institute Technical Report WIS-SG-92-250.
