---
title: "Statistical Estimation of Feeding Level from Field Data"
author: "Hans Ttito"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Statistical Estimation of Feeding Level from Field Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 7,
  fig.height = 5,
  message   = FALSE,
  warning   = FALSE
)
library(fb4package)
```

## Overview

A central challenge in fish bioenergetics is estimating the **proportion of
maximum consumption** (*p*) that a fish population actually achieves ŌĆö and
quantifying the uncertainty in that estimate. `fb4package` provides three
estimation strategies, each suited to a different data situation:

| Strategy | Data required | Output |
|---|---|---|
| `binary_search` | Single target weight | Point estimate of *p* |
| `mle` | Sample of final weights | *p*, SE, 95 % CI, AIC |
| `bootstrap` | Sample of final weights | *p*, bootstrap CI, distribution |

This vignette demonstrates all three on the same simulated dataset and compares
their outputs.

---

## 1. Shared model setup

We use a 100-day simulation for a generic salmonid with manually specified
parameters, so results are fully reproducible without the built-in database.

```{r shared-setup}
days <- 1:100

# Species parameters (Stewart & Ibarra 1991 ŌĆö Chinook salmon juvenile)
sp_params <- list(
  consumption = list(
    CEQ = 2, CA = 0.303, CB = -0.275,
    CQ = 3.0, CTO = 15.0, CTM = 20.0,
    CTL = 24.0, CK1 = 0.1, CK4 = 0.13
  ),
  respiration = list(
    REQ = 1, RA = 0.00264, RB = -0.217,
    RQ = 0.06818, RTO = 0.0234, RTM = 0.0,
    RTL = 25.0, RK1 = 1.0, RK4 = 0.13, RK5 = 0.0
  ),
  activity    = list(ACT = 1.0, BACT = 0.0),
  sda         = list(SDA = 0.172),
  egestion    = list(EGEQ = 1, FA = 0.212, FB = -0.222, FG = 0.631),
  excretion   = list(EXEQ = 1, UA = 0.0314, UB = 0.58, UG = -0.299),
  predator    = list(
    PREDEDEQ = 2,
    Alpha1 = 4500, Beta1 = 6.0,
    Alpha2 = 5500, Beta2 = 2.0,
    Cutoff = 250
  )
)

sp_info <- list(
  scientific_name = "Oncorhynchus tshawytscha",
  common_name     = "Chinook salmon",
  life_stage      = "juvenile"
)

# Seasonal temperature (┬░C)
temp_data <- data.frame(
  Day         = days,
  Temperature = round(8 + 6 * sin(2 * pi * (days - 20) / 365), 2)
)

# Two-prey diet with seasonal shift
alewife <- pmax(0, 0.65 + 0.20 * sin(2 * pi * (days - 30) / 365))
shrimp  <- pmax(0, 1 - alewife)
diet_props  <- data.frame(Day = days,
                           Alewife = round(alewife, 4),
                           Shrimp  = round(shrimp,  4))
prey_energy <- data.frame(Day = days, Alewife = 4900, Shrimp = 3200)

bio <- Bioenergetic(
  species_params      = sp_params,
  species_info        = sp_info,
  environmental_data  = list(temperature = temp_data),
  diet_data           = list(
    proportions = diet_props,
    prey_names  = c("Alewife", "Shrimp"),
    energies    = prey_energy
  ),
  simulation_settings = list(initial_weight = 10, duration = 100)
)
```

---

## 2. Simulated field data

We first run a simulation at a "true" *p* = 0.70 to obtain the deterministic
final weight, then draw 30 observed weights around that value with 6 % CV to
mimic field sampling variability.

```{r field-data}
# Reference simulation at true p
res_ref <- run_fb4(bio,
                   fit_to    = "p_value",
                   fit_value = 0.70,
                   strategy  = "direct_p_value",
                   verbose   = FALSE)

true_final_wt <- res_ref$summary$final_weight
cat(sprintf("True final weight at p = 0.70 : %.1f g\n", true_final_wt))

# Simulate 30 field observations
set.seed(2024)
n_obs       <- 30
obs_weights <- rnorm(n_obs,
                     mean = true_final_wt,
                     sd   = true_final_wt * 0.06)

cat(sprintf("Observed weights (n = %d) : mean = %.1f g  SD = %.1f g\n",
            n_obs, mean(obs_weights), sd(obs_weights)))
```

---

## 3. Strategy 1 ŌĆö Binary search on the mean weight

The simplest approach: find the *p* that produces a final weight equal to the
**sample mean** of the observed weights.

```{r binary-search}
res_bs <- run_fb4(bio,
                  fit_to    = "Weight",
                  fit_value = mean(obs_weights),
                  strategy  = "binary_search",
                  verbose   = FALSE)

cat(sprintf("Binary search  p = %.4f  (final weight = %.1f g)\n",
            res_bs$summary$p_value,
            res_bs$summary$final_weight))
```

This is fast and unambiguous, but provides **no information about uncertainty**.

---

## 4. Strategy 2 ŌĆö Maximum Likelihood Estimation

MLE assumes final weights follow a **log-normal distribution** around the
model-predicted weight.  It estimates *p* and the observation error (Žā)
simultaneously, returning standard errors and confidence intervals derived
from the Fisher information matrix.

```{r mle}
res_mle <- run_fb4(bio,
                   fit_to           = "Weight",
                   observed_weights = obs_weights,
                   strategy         = "mle",
                   estimate_sigma   = TRUE,
                   confidence_level = 0.95,
                   compute_profile  = FALSE,
                   verbose          = FALSE)

cat(sprintf("MLE  p = %.4f  SE = %.4f\n",
            res_mle$summary$p_estimate,
            res_mle$summary$p_se))
cat(sprintf("     95%% CI : [%.4f, %.4f]\n",
            res_mle$method_data$confidence_intervals$p_ci_lower,
            res_mle$method_data$confidence_intervals$p_ci_upper))
cat(sprintf("     sigma  = %.4f   AIC = %.2f   converged = %s\n",
            res_mle$summary$sigma_estimate,
            res_mle$method_data$aic,
            res_mle$summary$converged))
```

### Likelihood profile

Setting `compute_profile = TRUE` evaluates the log-likelihood across a grid
of *p* values, providing a visual check on the shape of the likelihood surface.

```{r mle-profile, cache=TRUE, fig.cap="Log-likelihood profile over p-values."}
res_mle_prof <- run_fb4(bio,
                         fit_to           = "Weight",
                         observed_weights = obs_weights,
                         strategy         = "mle",
                         estimate_sigma   = TRUE,
                         compute_profile  = TRUE,
                         profile_grid_size = 20,
                         verbose          = FALSE)

plot(res_mle_prof, type = "uncertainty")
```

---

## 5. Strategy 3 ŌĆö Bootstrap

Bootstrap resampling makes **no distributional assumptions**.  At each
iteration it resamples the observed weights with replacement, computes the
mean, and finds the *p* that matches it via optimisation.  The resulting
distribution of *p* estimates is used for inference.

```{r bootstrap, cache=TRUE}
res_boot <- run_fb4(bio,
                    fit_to           = "Weight",
                    observed_weights = obs_weights,
                    strategy         = "bootstrap",
                    n_bootstrap      = 200,
                    parallel         = FALSE,
                    confidence_level = 0.95,
                    verbose          = FALSE)

cat(sprintf("Bootstrap  p mean = %.4f  SD = %.4f\n",
            res_boot$summary$p_mean,
            res_boot$summary$p_sd))
cat(sprintf("           95%% CI : [%.4f, %.4f]\n",
            res_boot$method_data$confidence_intervals$p_ci_lower,
            res_boot$method_data$confidence_intervals$p_ci_upper))
cat(sprintf("           Success rate : %.1f%%\n",
            res_boot$method_data$bootstrap_info$success_rate * 100))
```

For long runs, activate parallel processing with `parallel = TRUE` and
`n_cores = parallel::detectCores() - 1`.

---

## 6. Comparing the three strategies

```{r comparison, echo=FALSE}
methods  <- c("True p", "Binary search", "MLE", "Bootstrap")
p_vals   <- c(0.700,
              res_bs$summary$p_value,
              res_mle$summary$p_estimate,
              res_boot$summary$p_mean)
ci_lower <- c(NA, NA,
              res_mle$method_data$confidence_intervals$p_ci_lower,
              res_boot$method_data$confidence_intervals$p_ci_lower)
ci_upper <- c(NA, NA,
              res_mle$method_data$confidence_intervals$p_ci_upper,
              res_boot$method_data$confidence_intervals$p_ci_upper)

result_tbl <- data.frame(
  Method   = methods,
  p_value  = round(p_vals,   4),
  CI_lower = round(ci_lower, 4),
  CI_upper = round(ci_upper, 4)
)

knitr::kable(result_tbl, caption = "Comparison of p estimates across strategies",
             col.names = c("Method", "p estimate", "95% CI lower", "95% CI upper"))
```

All three methods recover the true *p* Ōēł 0.70 closely.  MLE and Bootstrap
both produce **95 % confidence intervals** that include the true value, as
expected with 30 observations and 6 % CV.

---

## 7. When to use each strategy

### Binary search

Use when you have a **single target weight** (e.g., mean of a stock assessment
sample) and only need a point estimate.  It is the fastest option and is
appropriate when uncertainty is not of scientific interest.

### MLE

Use when you have a **sample of observed weights** and want:

- Asymptotic confidence intervals from the Fisher information matrix
- Model comparison via AIC (e.g., comparing diet or temperature scenarios)
- A formal likelihood-based framework for publication

Assumes log-normal errors; check this assumption by inspecting the residuals
from the final simulation.

### Bootstrap

Use when you want **non-parametric CIs** that do not rely on distributional
assumptions, or when the likelihood surface is irregular.  Bootstrap is
also preferable with small sample sizes (*n* < 15) where asymptotic MLE CIs
may be unreliable.

For both statistical methods, at least **5 observations** are required
(`bootstrap` requires Ōēź 5; `mle` requires Ōēź 3).

---

## References

Deslauriers D, Chipps SR, Breck JE, Rice JA, Madenjian CP (2017). Fish
Bioenergetics 4.0: An R-Based Modeling Application. *Fisheries* 42(11):586ŌĆō596.
<https://doi.org/10.1080/03632415.2017.1377558>

Efron B, Tibshirani RJ (1993). *An Introduction to the Bootstrap*.
Chapman and Hall/CRC, New York.
